Estimating the magnetization distribution within rectangular rock samples

### 2016-06-15 manuscript/manuscript.tex

\documentclass[draft,gc]{agutex}
%\documentclass[galley,gc]{agutex}

@@

Based on the estimation of the total (bulk) remanent magnetization, the standard 
paleomagnetic techniques have been used for more than 40 years in studies of rock magnetism, 
magnetostratigraphic and paleogeographic reconstructions \citep{voo1993}. These techniques 
are well established and are applied successfully in almost all paleomagnetic studies. 
However, this classical method has constraints in studies that need high spatial or 
temporal resolution, such as the secular variation's investigation. In this context, new 
methods have been developed over the last decades to get around these limitations, in special, 
the Scanning Magnetic Microscopy (SMM) techniques \citep{oda2011, fu2014}. Different from 
standard paleomagnetic magnetometers, the SMM devices produce high-resolution maps of the 
vertical component of the magnetic induction field on a planar surface located over the 
sample. Usually, the scanning magnetic microscopes based on superconducting quantum 
interference device (SQUID) sensors achieve the best field sensitivity for geoscientific 
research. However, these devices are very expensive to build/operate and require specific 
technologies to keep the sensor at cryogenic temperatures and also present a limited 
sensor-to-sample distance $\approx 100$~$\mu$m \citep{baudenbacher2003, fong2005}. 
Despite of these technical difficulties, SQUID sensors have been successfully applied 
in several paleomagnetic studies \citep{gattacceca2006,weiss2007b, oda2011, fu2014}. 
To avoid the problems related with the SQUID complexity, numerous studies have been 
attempted to develop alternative low-cost high-performance SMM instruments based on 
magneto-impedance (MI) \citep{uehara2007, uehara2008}, giant magnetoresistance (GMR) 
\citep{hankard2009} and magnetic tunnel junction (MTJ)\citep{lima2014} sensors. 
Compared with SQUID's, these alternative sensors are less sensitive. On the other 
hand, they are generally easier to operate, operate at room temperature, 
may achieve higher spatial resolution and also a reduction in the sensor-to-sample 
distance.

@@

The parameter vector $\mathbf{m} = \mathbf{m}^{\dagger}$ minimizing 
the goal function (equation \ref{eq:goal-function}) is given by
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} \right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: ,
\label{eq:estimate}
\end{equation}
where $\mathbf{M}$ is the partitioned matrix defined in equation 
\ref{eq:sensitivity-matrix}.
We consider that the estimated parameter vector $\mathbf{m}^{\dagger}$
given by equation \ref{eq:estimate} approximates the real magnetization 
distribution within the rectangular rock sample. However, solutions given by the equation \ref{eq:estimate} can present non-uniqueness and instability. To overcome this ill- conditioning problem we need to impose a \textit{priori} information. As mentioned before, this procedure is generally called regularization. Here we imposed a constraint using ''Smoothness" regularization, generally called first-order Tikhonov regularization. Then equation \ref{eq:estimate} is now equal to 
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} + \mu_0 \,  \mathbf{R}^T \mathbf{R}\right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: ,
\label{eq:estimate_regul}
\end{equation}
where $\mu_0$ is a positive number called parameter regularization and $\mathbf{R}$ is a matrix with elements $-1$ and $1$. For example, if we have a sample formed by two prisms ($P=2$), the matrix $\mathbf{R}$ can be written as

@@

@@

\begin{thebibliography}{39}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

@@

@@

>>>>

%\documentclass[draft,gc]{agutex}
\documentclass[galley,gc]{agutex}

@@

Based on the estimation of the total (bulk) remanent magnetization, the standard 
paleomagnetic techniques have been used for more than 40 years in studies of rock magnetism, 
magnetostratigraphic and paleogeographic reconstructions \citep{voo1993}. These techniques 
are well established and are applied successfully in almost all paleomagnetic studies. 
However, this classical method has constraints in studies that need high spatial or 
temporal resolution, such as the secular variation's investigation. In this context, new 
methods have been developed over the last decades to get around these limitations, in special, 
the Scanning Magnetic Microscopy (SMM) techniques \citep{oda2011, fu2014}. Different from 
standard paleomagnetic magnetometers, the SMM devices produce high-resolution maps of the 
vertical component of the magnetic induction field on a planar surface located over the 
sample. Usually, the scanning magnetic microscopes based on superconducting quantum 
interference device (SQUID) sensors achieve the best field sensitivity for geoscientific 
research. However, these devices are very expensive to build/operate and require specific 
technologies to keep the sensor at cryogenic temperatures and also present a limited 
sensor-to-sample distance $\approx 100$~$\mu$m \citep{baudenbacher2003, fong2005}. 
Despite of these technical difficulties, SQUID sensors have been successfully applied 
in several paleomagnetic studies \citep{gattacceca2006,weiss2007b, oda2011, fu2014}. 
To avoid the problems related with the SQUID complexity, numerous studies have been 
attempted to develop alternative low-cost high-performance SMM instruments based on 
magneto-impedance (MI) \citep{uehara2007, uehara2008}, giant magnetoresistance (GMR) 
\citep{hankard2009} and magnetic tunnel junction (MTJ) \citep{lima2014} sensors. 
Compared with SQUID's, these alternative sensors are less sensitive. On the other 
hand, they are generally easier to operate, operate at room temperature, 
may achieve higher spatial resolution and also a reduction in the sensor-to-sample 
distance.

@@

The parameter vector $\mathbf{m} = \mathbf{m}^{\dagger}$ minimizing 
the goal function (equation \ref{eq:goal-function}) is given by
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} \right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: ,
\label{eq:estimate}
\end{equation}
where $\mathbf{M}$ is the partitioned matrix defined in equation 
\ref{eq:sensitivity-matrix}.
We consider that the estimated parameter vector $\mathbf{m}^{\dagger}$
given by equation \ref{eq:estimate} approximates the real magnetization 
distribution within the rectangular rock sample. However, solutions given by the equation \ref{eq:estimate} can present non-uniqueness and instability. To overcome this ill- conditioning problem we need to impose a \textit{priori} information. As mentioned before, this procedure is generally called regularization. Here we imposed a constraint using ``Smoothness" regularization, generally called first-order Tikhonov regularization. Then equation \ref{eq:estimate} is now equal to 
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} + \mu_0 \,  \mathbf{R}^T \mathbf{R}\right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: ,
\label{eq:estimate_regul}
\end{equation}
where $\mu_0$ is a positive number called parameter regularization and $\mathbf{R}$ is a matrix with elements $-1$ and $1$. For example, if we have a sample formed by two prisms ($P=2$), the matrix $\mathbf{R}$ can be written as

@@

\subsection{Sensor active area}
\label{subsec:Sensor active area}

Magnetic sensors do not measure the magnetic induction 
field at a point. They measure instead the field averaged 
over the sensor active area (or volume), which is equivalent to a 
convolution of the magnetic induction field with 
the sensor answer function \citep{roth1989, egli2000, lima2013}.
The sensor-to-sample distance and the effective area (or volume) 
are important limiting factors determining the spatial resolution 
of magnetic sensors.

The magnetic sensors used in paleomagnetic applications have
different active area (or volume). 
\citet{uehara2007, uehara2008} presented a magneto-impedance (MI) 
sensor whose cylindrical active volume has $36\mu$m in diameter and 
$5500\mu$m in length, which limits its spatial resolution
if compared with the giant magnetoresistance 
(GMR) sensor presented by \citet{hankard2009} and the magnetic 
tunnel junction (MTJ) sensor presented by \citet{lima2014}, for example.
These two sensors are considered planar and have active areas
of, respectively, $9\mu$m~$\times$~$36\mu$m and 
$\approx 4\mu$m~$\times$~$2\mu$m.
Finally, \citet{fong2005} presented a low-transition-temperature 
superconducting quantum interference device (low T$_{c}$ SQUID) sensor
which is also considered planar and has an active area with
diameter of $80\mu$m.
This is currently one of the most powerful magnetic sensors.

In the following, we present an analysis of the averaging effect produced
by the finite size of sensors. We limit our study to squared-planar 
sensors which are parallel to the sample surface.
Let us first consider the magnetic induction component $\tilde{b}(x,y)$ that is 
parallel to the sample surface and is threaded
through a planar sensing area parallel to the sample
as the following two-dimensional convolution
\begin{equation}
\tilde{b}(x, y) = \int \limits_{-\infty}^{\infty} \int \limits_{-\infty}^{\infty}
w(\eta, \zeta) \; b(x - \eta, y - \zeta) \, d\eta d\zeta \: ,
\label{eq:b-mag-flux}
\end{equation}
where $w(\eta, \zeta)$ is the sensor answer function \citep{egli2000}. 
Here, we do not consider the
dependence with the sensor-to-sample distance $h$ 
(Figure \ref{fig:b-mag-flux-approx}a).
For convenience, we consider a square sensor having an active area $A$
(light grey square in Figure \ref{fig:b-mag-flux-approx}a).
In this case, the sensor answer function $w(\eta, \zeta)$ 
is given by \citep{egli2000}
\begin{equation}
w(x, y) = \begin{cases}
\begin{split}
1/A &\: , \quad (x, y) \in A \\
0 &\: , \quad (x, y) \notin A
\end{split}
\end{cases} \: .
\label{eq:w-function}
\end{equation}
By substituting this sensor answer function into the
magnetic flux integral (equation \ref{eq:b-mag-flux})
we obtain
\begin{equation}
\tilde{b}(x, y) = \frac{1}{A} 
\int \limits_{-\infty}^{\infty} \int \limits_{-\infty}^{\infty}
b(x - \eta, y - \zeta) \, d\eta d\zeta \: .
\label{eq:b-mag-flux2}
\end{equation}
Equation \ref{eq:b-mag-flux2} has an analytical solution
in the wavenumber domain \citep{egli2000}.
However, we opted for using an approximated solution in the
space domain. This approximation consists in discretizing the 
integral \ref{eq:b-mag-flux2} by using a midpoint rule on 
a regular grid of $Q_{x} \times Q_{y}$ points $(x_{j}, y_{j}) \in A$,
$j = 1, ..., n$, where $Q = Q_{x} \, Q_{y}$.
The approximated integral is given by
\begin{equation}
\tilde{b}_{i} \approx \frac{\Delta A}{A} \sum \limits_{j = 1}^{Q} b_{j} \: ,
\label{eq:b-mag-flux-approx}
\end{equation}
where $\tilde{b}_{i} \equiv \tilde{b}(x_{i}, y_{i})$, $(x_{i}, y_{i})$ is
the point located at the center of the sensor active area $A$
(black circle in Figure \ref{fig:b-mag-flux-approx}b) and 
$b_{j}$ is the magnetic induction component parallel to the
sample surface and evaluated at the point $(x_{j}, y_{j})$
(open circle in Figure \ref{fig:b-mag-flux-approx}c),
$j = 1, ..., Q$.
According to the discretization represented in Figure 
\ref{fig:b-mag-flux-approx}b, the active sensor area is
divided into a regular grid of square cells with area
$\Delta A$ (Figure \ref{fig:b-mag-flux-approx}c). 
Hence, by noting that $A = Q \, \Delta A$, we get
\begin{equation}
\tilde{b}_{i} \approx \frac{1}{Q} \sum \limits_{j = 1}^{Q} b_{j} \: .
\label{eq:b-mag-flux-approx2}
\end{equation}
Equation \ref{eq:b-mag-flux-approx2} shows that the magnetic induction
$\tilde{b}_{i}$ at the observation point $(x_{i}, y_{i})$ is 
approximately given by the magnetic induction
averaged over the active area $A$ of the sensor.

@@

\begin{thebibliography}{40}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

@@

\bibitem[{\textit{Roth et~al.}(1989)\textit{Roth, Sepulveda, and
  Wikswo}}]{roth1989}
Roth, B.~J., N.~G. Sepulveda, and J.~P. Wikswo (1989), Using a magnetometer to
  image a two-dimensional current distribution, \textit{J. Appl. Phys.
  (Melville, NY, U. S.)}, \textit{65}(1), 361--372, \doi{10.1063/1.342549}.

@@

 \begin{figure}
 \noindent \includegraphics[width=20pc]{Figs/Fig4b_HQ.pdf}
 \caption{Schematic representation of the integral \ref{eq:Phi-mag-flux2}.
 (a) Magnetic sensor with active area $A$ (light grey square) located at
 a distance $h$ of the sample (grey prism). (b) Detailed representation
 of the sensor active area, which is centered at the observation point
 $(x_{i}, y_{i})$ (black circle). The square active area is discretized into 
 a regular grid of $Q$ square cells (limited by the dashed lines). 
 (c) jth cell with area $\Delta A$ and center at the point $(x_{j}, y_{j})$
 (open circle), $j = 1, ..., Q$.}
 \label{fig:b-mag-flux-approx}
 \end{figure}
